---
title: "RepData_Project1"
author: "Daniela Galatro"
date: "December 21, 2018"
output: html_document
---

##Assignment Steps
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

##Step 1
###Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
'Load the data'
setwd("C:/Users/CocoCuchi/Desktop/JohnHopkins/Reproducible/Project1/repdata_data_activity")
'No processing of the data is required; format is suitable for analysis'
activity <-read.csv("activity.csv")
```

##Step 2
###Histogram of the total number of steps taken each day
```{r, echo = TRUE}
'Estimate number of steps taken per day, removing NAs'
stepsday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
'Plotting histogram'
png("plot1.png")
  hist(stepsday$steps, xlab = "Steps per Day", ylim=c(0,30),main = "Total number of steps taken per day", labels=TRUE, col = "blue4")
dev.off()

```

##Step 3
###Mean and median number of steps taken each day
```{r, echo = TRUE}
mean.steps <- format(mean(stepsday$steps),digits=1)
median.steps <- format(median(stepsday$steps),digits=1)
'Report mean and media of number of steps taken each interval'
mean.steps
median.steps
```

##Step 4
###Time series plot of the average number of steps taken
```{r, echo = TRUE}
steps.mean <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
png("plot4.png")
  plot(steps.mean$interval, steps.mean$steps, type = "b", col = "blue4", xlab = "Intervals", ylab = "Total steps per interval", main = "Number of steps per interval (averaged)")
dev.off()
```
##Step 5
###The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}
max.steps <-max(steps.mean$steps)
max.interval <-steps.mean$interval[which(steps.mean$steps == max.steps)]
max.steps=round(max.steps,0)
'Report values'
max.steps
max.interval
```
##Step 6
###Code to describe and show a strategy for imputing missing data
```{r, echo = TRUE}
Steps.Interval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
# Split activity data per interval and fill missing data
splitdata <- split(activity, activity$interval)
for(i in 1:length(splitdata)){
    splitdata[[i]]$steps[is.na(splitdata[[i]]$steps)] <- Steps.Interval[i]
}
data.imputed <- do.call("rbind", splitdata)
data.imputed  <- data.imputed[order(data.imputed$date) ,]
```

##Step 7
###Histogram of the total number of steps taken each day after missing values are imputed
'Plotting histogram'
```{r, echo = TRUE}
png("plot6.png")
  Steps.imputed <- tapply(data.imputed$steps, data.imputed$date, sum)
  hist(Steps.imputed, xlab = "Steps per Day", ylim=c(0,40), main = "Total number of steps taken per day (Imputed data)",labels=TRUE, col = "blue4")
dev.off()
'Report mean and median for total number of steps taken per day'
Mean.imputed <- mean(Steps.imputed, na.rm = TRUE)
Median.imputed <- median(Steps.imputed, na.rm = TRUE)
Mean.imputed
Median.imputed
```

##Step 8
###Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo = TRUE}
'New factor variable in the dataset for weekday and weekend'
data.imputed$day <- ifelse(weekdays(as.Date(data.imputed$date)) == "Saturday" | weekdays(as.Date(data.imputed$date)) == "Sunday", "weekend", "weekday")
'Average steps per interval for weekdays'
StepsPInt.weekday <- tapply(data.imputed[data.imputed$day == "weekday" ,]$steps, data.imputed[data.imputed$day == "weekday" ,]$interval, mean, na.rm = TRUE)
'Average steps for internal for weekends'
StepsPInt.weekend <- tapply(data.imputed[data.imputed$day == "weekend" ,]$steps, data.imputed[data.imputed$day == "weekend" ,]$interval, mean, na.rm = TRUE)
'Two Panels Plot'
png("plot8.png")
par(mfrow=c(1,2))
'Plot weekday activity'
p1 <- plot(as.numeric(names(StepsPInt.weekday)), StepsPInt.weekday, xlab = "Interval", ylab = "Steps", 
     ylim = c(0,250), main = "Activity Pattern - Weekdays", col="blue4", type = "b")
'Plot weekend activity'
p2 <- plot(as.numeric(names(StepsPInt.weekend)), StepsPInt.weekend, xlab = "Interval", ylab = "Steps", 
     ylim = c(0,250), main = "Activity Pattern - weekends", col="red4", type = "b")
dev.off()
```